---
title: "Survey on ethics regarding AI-generated art"
subtitle: "Data Acquisition and Survey Methods"
output: html_notebook
author: "Nikolaus Czernin, Aliakseyeu Dzimitry"
---

# Requirements (delete)

Survey data analysis, report & presentation:

-   **Exploratory Data Analysis**: Choose appropriate methods to explore the categorical and quantitative features in your data set ([one visualization per research question]{.underline}). Briefly explain the methods and comment on your findings.

-   **Descriptive Inference**: Compute suitable [summary statistics]{.underline} and display the [results in a table]{.underline}.

-   **Analytic Inference**: Select an appropriate test statistic to [test your hypothesis]{.underline}. Summarize the results and explain how your findings are connected to your research question.

\
Guidelines for the report:

-   use RMarkdown

-   include a [table of contents]{.underline}

-   briefly summarize the functions used

-   display all [important]{.underline} code snippets in the .pdf

-   document your findings

Submit both the [.rmd and .pdf]{.underline} files via TUWEL.

```{r, include=FALSE}
# @Nate: Guidlines say to only include important code snippets
# I would argue that noone will read this snipped
# therefore I set include to FALSE
# feel free to change it if you disagree
library(tidyverse)
library(ggpubr)
library(forcats)
```

```{r}
ans <- read.delim("answers.csv", sep = ";")
names(ans) <- c("gender", "age", "program", "a1", "a2", "a3")
ans <-
  ans %>% 
  # rename the answer cols and transform them 
  mutate(
    revenue.deserved = str_extract(a1, "[:digit:]*") %>% as.numeric(),
    is.artist = a2 %>% as.factor(),
    would.boycott = a3 %>% as.factor()
    ) %>% 
  select(-a1, -a2, -a3)
ans
ans %>% summary()

ans %>% select(is.artist)
# recode the factor levels of the longer answers
levels(ans$is.artist) <- c("I'm an artist", "I enjoy creating art", 
                                "I created some art, not passionately", "I do not create")
ans %>% select(is.artist)
```

# Exploratory analysis

```{r}
p1 <- ans %>% 
  ggplot(aes(revenue.deserved)) +
  geom_histogram(bins=10)
p2 <- ans %>% 
  ggplot(aes(revenue.deserved)) +
  geom_boxplot() 
ggarrange(p1, p2, nrow=2)

```

```{r}
ans %>% 
  filter(age<30) %>% 
  ggplot(aes(age, revenue.deserved)) +
  geom_point()
```

```{r, fig.width=7, fig.height=7}

ans %>% 
  drop_na() %>% 
  group_by(is.artist) %>% 
  summarise(
    `Average percentage of revenue granted to an AI-trained-on artist` = mean(revenue.deserved),
    `Total number of participants`=n()
  ) %>% 
  pivot_longer(c(`Total number of participants`,
                 `Average percentage of revenue granted to an AI-trained-on artist`
                 )) %>% 
  ggplot(aes(x=is.artist, y=value)) +
  geom_bar(stat="identity") +
    facet_wrap(~name, ncol = 1, scale="free") +
  xlab("Participants grouped by their relationship to creating art") +
  ylab(NULL)

```

