---
title: "Survey on ethics regarding AI-generated art"
subtitle: "Data Acquisition and Survey Methods"
output: html_notebook
author: "Nikolaus Czernin, Aliakseyeu Dzimitry"
---

# Requirements (delete)

Survey data analysis, report & presentation:

-   **Exploratory Data Analysis**: Choose appropriate methods to explore the categorical and quantitative features in your data set ([one visualization per research question]{.underline}). Briefly explain the methods and comment on your findings.

-   **Descriptive Inference**: Compute suitable [summary statistics]{.underline} and display the [results in a table]{.underline}.

-   **Analytic Inference**: Select an appropriate test statistic to [test your hypothesis]{.underline}. Summarize the results and explain how your findings are connected to your research question.

\
Guidelines for the report:

-   use RMarkdown

-   include a [table of contents]{.underline}

-   briefly summarize the functions used

-   display all [important]{.underline} code snippets in the .pdf

-   document your findings

Submit both the [.rmd and .pdf]{.underline} files via TUWEL.

```{r, include=FALSE}
# @Nate: Guidlines say to only include important code snippets
# I would argue that noone will read this snipped
# therefore I set include to FALSE
# feel free to change it if you disagree
library(tidyverse)
library(ggpubr)

```

```{r}
ans <- read.delim("answers.csv", sep = ";")
names(ans) <- c("gender", "age", "program", "a1", "a2", "a3")
ans <-
  ans %>% 
  mutate(
    a1 = str_extract(a1, "[:digit:]*") %>% as.numeric(),
    a2 = a2 %>% as.factor(),
    a3 = a3 %>% as.factor()
    ) 
ans
ans %>% summary()
```

# Exploratory analysis

```{r}
p1 <- ans %>% 
  ggplot(aes(a1)) +
  geom_histogram(bins=10)
p2 <- ans %>% 
  ggplot(aes(a1)) +
  geom_boxplot() 
ggarrange(p1, p2, nrow=2)

```

```{r}
ans %>% 
  filter(age<30) %>% 
  ggplot(aes(age, a1)) +
  geom_point()
```
