---
title: "Survey on ethics regarding AI-generated art"
subtitle: "Data Acquisition and Survey Methods"
output: html_notebook
author: "Nikolaus Czernin, Aliakseyeu Dzimitry"
---

# Requirements (delete)

Survey data analysis, report & presentation:

-   **Exploratory Data Analysis**: Choose appropriate methods to explore the categorical and quantitative features in your data set ([one visualization per research question]{.underline}). Briefly explain the methods and comment on your findings.

-   **Descriptive Inference**: Compute suitable [summary statistics]{.underline} and display the [results in a table]{.underline}.

-   **Analytic Inference**: Select an appropriate test statistic to [test your hypothesis]{.underline}. Summarize the results and explain how your findings are connected to your research question.

\
Guidelines for the report:

-   use RMarkdown

-   include a [table of contents]{.underline}

-   briefly summarize the functions used

-   display all [important]{.underline} code snippets in the .pdf

-   document your findings

Submit both the [.rmd and .pdf]{.underline} files via TUWEL.

```{r, include=FALSE}
library(tidyverse)
library(ggpubr)
library(forcats)
library(knitr)


library("fabricatr")
```

```{r}
ans <- read.delim("answers.csv", sep = ";")
names(ans) <- c("gender", "age", "program", "a1", "a2", "a3")

ans <- ans %>% 
  # rename the answer cols and transform them 
  mutate(
    revenue.deserved = str_extract(a1, "[:digit:]*") %>% as.numeric(),
    is.artist = a2 %>% as.factor(),
    would.boycott = a3 %>% as.factor()
    ) %>% 
  select(-a1, -a2, -a3)

# recode the factor levels of the longer answers
levels(ans$is.artist) <- c("I'm an artist", "I enjoy creating art",
                                "I created some art,\nnot passionately", "I do not create art")
# reverse the levels
ans$is.artist <- ans$is.artist %>% fct_rev()
# rearrange the factor levels of the text answers
# levels(ans$is.artist) = c(4, 3, 2, 1) # the levels for is.artist arrange themselves perfectly somehow; 1 is no art, 4 is max art
# levels(ans$would.boycott) = c(0, 1, 2)

# fill the single NA value in the revenue.deserved column with the median value
ans <- ans %>% mutate(across(revenue.deserved, ~replace_na(., median(., na.rm=TRUE))))


# assigning short variable names to the questions
q1_varname <- "Percentage of generated revenue granted to trained on artist"
q2_varname <- "Respondents' relationships to creating art"
q3_varname <- "Respondents' propensity to boycott unpaid AI-generated art training"
```
```{r, echo=F}
ans %>% head()
ans %>% summary()

```


## Research hypothesis 1

*Artists/authors would credit other artists/authors to more money, if the model were trained on them, than non-artists/non-authors.*

### Exploring the nature of the data/responses

```{r, echo=F}

ans %>% 
  group_by(is.artist) %>% 
  summarise(
    n=n(),
    `Mean age`=mean(age, na.rm=T) %>% round(1)
  ) %>% 
  ggplot(aes(x=is.artist, y=n, fill=`Mean age`)) +
  geom_bar(stat="identity") +
  geom_text(aes(label=n, x=is.artist, y=n+1.5)) +
  theme_minimal() +
  ylab(NULL) +
  xlab(NULL) +
  ggtitle(q2_varname)

# Histogram
q1_hist <- ans %>% 
  filter(!is.na(revenue.deserved)) %>% 
  ggplot(aes(revenue.deserved)) +
  geom_histogram(bins=10) +
  ggtitle("Reported proceeds allocated to the artist") +
  theme_minimal() +
  ylab(NULL) +
  xlab(NULL)

# Boxplot overall
q1_boxplot <- ans %>% 
  filter(!is.na(revenue.deserved)) %>% 
  ggplot(aes(revenue.deserved)) +
  theme_minimal() +
  geom_boxplot() +
  theme(
    axis.text.y=element_blank(),axis.ticks=element_blank(), 
  ) +
  xlab(q1_varname)
# put them on top of each other
ggarrange(q1_hist, q1_boxplot, nrow=2) %>% print()

# Boxplot per level of questison 2: are you an artist youself?
ans %>% 
  filter(!is.na(revenue.deserved)) %>% 
  ggplot(aes(x=revenue.deserved, y=is.artist)) +
  geom_boxplot() +
  theme_minimal() +
  xlab(q1_varname) +
  ylab(NULL) +
  ggtitle("Reported proceeds allocated to the artist by relationship to creating art") 

```

### Independence testing.

Null hypothesis: If a generative model was trained entirely on the works of one artist, then other artists would credit a higher percentage of the proceeds of said model to the original artists

```{r}
k <- 5
# cut the revenue.deserved values into k levels: low to high 
ans$revenue.deserved.cat = split_quantile(ans$revenue.deserved, type=k)  
table(ans$revenue.deserved.cat, ans$is.artist) %>% 
  print() %>% 
  chisq.test()

# what if we replicated that data with the same results?
table(ans$revenue.deserved.cat, ans$is.artist) %>% 
  chisq.test(simulate.p.value = TRUE)

```

We conducted chi-squared tests to examine the independence of the percentage of 
revenue a respondent would grant an artist, that the generative model had been trained on
and the respondent's relationship to creating art themselves. 
We recoded the granted.revenue variable to the quintiles. 
To correct for the low sample size, we run the test with replicates. 
The resulting p-values were ~0.7 and ~0.8, which exceed the conventional significance level of 0.05. 
As a result, there is insufficient evidence to reject the null hypothesis of independence, 
even when using replicates. A bigger pool of respondents may have yielded different 
results, after all, only 6 of the respondents claimed to be somewhat passionate 
about creating art. 



```{r, include=FALSE}

#### Regression
# For good measure we compute a linear model. 
# As expected, it is completely useless, we are nowhere near being able
# to reject the Null-Hypothesis of any of the factors being non-zero. 

rs1 <- ans %>% 
  drop_na() %>% 
  select(revenue.deserved, is.artist) %>% 
  mutate(x=row_number(), y=1) %>% 
  pivot_wider(id_cols=c(x, revenue.deserved), names_from = is.artist, values_from =y ) %>% 
  select(-x) %>% 
  mutate_all(function(x) replace_na(x, 0))

linear_model <- lm(revenue.deserved ~ ., rs1)
summary(linear_model)

```



## Research hypothesis 2

*Artists/authors are, compared to non-artists/non-authors, more inclined to boycott art that was created by a model that trained on an artist, if the artist doesn't get any money out of it*


```{r, echo=F, fig.width=9}

boycotts <- ans %>% 
  group_by(would.boycott, is.artist) %>% 
  summarise(n=n()) %>% 
  group_by(would.boycott) %>% 
  mutate(total=sum(n)) 

boycotts %>% 
  ggplot(aes(x=would.boycott,y=n, fill=is.artist)) +
    geom_bar(stat = "identity") +
    geom_text(aes(label=total, x=would.boycott, y=total+1)) +
    scale_fill_brewer(palette="Paired") +
    theme_minimal() +
    labs(fill=q2_varname) +
    theme(legend.position = "bottom") +
    ggtitle(q3_varname) +
    xlab(NULL) +
    ylab(NULL) 

```

Turns out, exactly half of the respondents would consider boycotting a piece of AI-generated art, 
if the trained-on artist was not compensated for his contribution. Some of them only 
if they already really liked the artist.  
  
Maybe more surprisingly, the two respondents claiming to be artists themselves 
answered not to be wanting to boycott such art. 

```{r, fig.width=10}
# step 1

ggplot(ans, aes(y=would.boycott, fill=is.artist)) + 
  geom_bar() + 
    scale_fill_brewer(palette="Paired") +
  labs(y = NULL, x=NULL,
       title=q3_varname %>% paste( "by level of engagement with art"),
       fill=q2_varname
       ) + 
  facet_wrap(~ is.artist)
```


### Independence testing.

Null hypothesis: If the artist the generative model was trained on were not granted any of the revenue generated by the art, other artists would be more likely to boycott that piece of art.  

From looking at the above barplots, we cannot make out a clear pattern of who is 
more likely to boycott art visually.


```{r, fig.width=10}

# cut the revenue.deserved values into k levels: low to high 
table(ans$would.boycott, ans$is.artist) %>% 
  print() %>% 
  chisq.test()

# what if we replicated that data with the same results?
table(ans$would.boycott, ans$is.artist) %>% 
  chisq.test(simulate.p.value = TRUE)


```

TODO: 
write up summary

```{r, fig.width=10}

# the reason for first converting to str then to int is because I gave the factors pre-determined int values
is.artist = as.numeric(as.character(ans$is.artist))
would.boycott = as.numeric(as.character(ans$would.boycott))

cor.test(is.artist, would.boycott, method = "spearman")

```


## Research hypothesis 3
*There is no statistical connection between people who would grant artists/authors more money (question 1) and people who would boycott an AI-generated piece of art, that the trained-on artist did not make any money from, meaning that respondents, while they may agree with the ethical worries of artists regarding AI-generated art, they do not act upon those values.*

### Exploring the data

```{r, fig.width=9}
ans %>%   
  ggplot(aes(x=revenue.deserved, y=would.boycott)) +
    geom_boxplot() +
    theme_minimal() +
    ylab(NULL) +
  ggtitle(q1_varname %>% paste("\nby", q3_varname)) +
  xlab(q1_varname)

```

Looking at the paralell boxplots, we can clearly tell that th