---
title: "Survey on ethics regarding AI-generated art"
output:
  pdf_document:
    toc: true
    toc_depth: 2
---


subtitle: "Data Acquisition and Survey Methods"\
author: "Nikolaus Czernin, Aliakseyeu Dzimitry"


```{r, include=FALSE}
library(tidyverse)
library(forcats)
library(fabricatr)
```

```{r}
ans <- read.delim("answers.csv", sep = ";")
names(ans) <- c("gender", "age", "program", "a1", "a2", "a3")

ans <- ans %>% 
  # rename the answer cols and transform them 
  mutate(
    revenue.deserved = str_extract(a1, "[:digit:]*") %>% as.numeric(),
    is.artist = a2 %>% as.factor(),
    would.boycott = a3 %>% as.factor()
    ) %>% 
  select(-a1, -a2, -a3)

# recode the factor levels of the longer answers
levels(ans$is.artist) <- c("I'm an artist", "I enjoy creating art",
                                "I created some art,\nnot passionately", "I do not create art")
# reverse the levels
ans$is.artist <- ans$is.artist %>% fct_rev()
# rearrange the factor levels of the text answers
# levels(ans$is.artist) = c(4, 3, 2, 1) # the levels for is.artist arrange themselves perfectly somehow; 1 is no art, 4 is max art
# levels(ans$would.boycott) = c(0, 1, 2)

# fill the single NA value in the revenue.deserved column with the median value
ans <- ans %>% mutate(across(revenue.deserved, ~replace_na(., median(., na.rm=TRUE))))


# assigning short variable names to the questions
q1_varname <- "Percentage of generated revenue granted to trained on artist"
q2_varname <- "Respondents' relationships to creating art"
q3_varname <- "Respondents' propensity to boycott unpaid AI-generated art training"
```

```{r, echo=F}
ans %>% head()
ans %>% summary()
```

# Research hypothesis 1

*Artists/authors would credit other artists/authors to more money, if the model were trained on them, than non-artists/non-authors.*

## Exploring the nature of the data/responses

```{r, echo=F}

ans %>% 
  group_by(is.artist) %>% 
  summarise(
    n=n(),
    `Mean age`=mean(age, na.rm=T) %>% round(1)
  ) %>% 
  ggplot(aes(x=is.artist, y=n, fill=`Mean age`)) +
  geom_bar(stat="identity") +
  geom_text(aes(label=n, x=is.artist, y=n+1.5)) +
  theme_minimal() +
  ylab(NULL) +
  xlab(NULL) +
  ggtitle(q2_varname)

# Histogram
q1_hist <- ans %>% 
  filter(!is.na(revenue.deserved)) %>% 
  ggplot(aes(revenue.deserved)) +
  geom_histogram(bins=10) +
  ggtitle("Reported proceeds allocated to the artist") +
  theme_minimal() +
  ylab(NULL) +
  xlab(NULL)

# Boxplot overall
q1_boxplot <- ans %>% 
  filter(!is.na(revenue.deserved)) %>% 
  ggplot(aes(revenue.deserved)) +
  theme_minimal() +
  geom_boxplot() +
  theme(
    axis.text.y=element_blank(),axis.ticks=element_blank(), 
  ) +
  xlab(q1_varname)
q1_hist %>% print()
q1_boxplot %>% print()

# Boxplot per level of questison 2: are you an artist youself?
ans %>% 
  filter(!is.na(revenue.deserved)) %>% 
  ggplot(aes(x=revenue.deserved, y=is.artist)) +
  geom_boxplot() +
  theme_minimal() +
  xlab(q1_varname) +
  ylab(NULL) +
  ggtitle("Reported proceeds allocated to the artist by relationship to creating art") 

```

## Independence testing.

*Hypothesis: If a generative model was trained entirely on the works of one artist, then other artists would credit a higher percentage of the proceeds of said model to the original artists*
```{r}
k <- 5
# cut the revenue.deserved values into k levels: low to high 
ans$revenue.deserved.cat = split_quantile(ans$revenue.deserved, type=k)  
table(ans$revenue.deserved.cat, ans$is.artist) %>% 
  print() %>% 
  chisq.test()

# what if we replicated that data with the same results?
table(ans$revenue.deserved.cat, ans$is.artist) %>% 
  chisq.test(simulate.p.value = TRUE)

```

We conducted chi-squared tests to examine the independence of the percentage of revenue a respondent would grant an artist, that the generative model had been trained on and the respondent's relationship to creating art themselves. We recoded the granted.revenue variable to the quintiles. To correct for the low sample size, we run the test with replicates. The resulting p-values were \~0.7 and \~0.8, which exceed the conventional significance level of 0.05. As a result, there is insufficient evidence to reject the null hypothesis of independence, even when using replicates. A bigger pool of respondents may have yielded different results, after all, only 6 of the respondents claimed to be somewhat passionate about creating art.

```{r, include=FALSE}

#### Regression
# For good measure we compute a linear model. 
# As expected, it is completely useless, we are nowhere near being able
# to reject the Null-Hypothesis of any of the factors being non-zero. 

rs1 <- ans %>% 
  drop_na() %>% 
  select(revenue.deserved, is.artist) %>% 
  mutate(x=row_number(), y=1) %>% 
  pivot_wider(id_cols=c(x, revenue.deserved), names_from = is.artist, values_from =y ) %>% 
  select(-x) %>% 
  mutate_all(function(x) replace_na(x, 0))

linear_model <- lm(revenue.deserved ~ ., rs1)
summary(linear_model)

```

# Research hypothesis 2

*Artists/authors are, compared to non-artists/non-authors, more inclined to boycott art that was created by a model that trained on an artist, if the artist doesn't get any money out of it*

```{r, echo=F, fig.width=9}

boycotts <- ans %>% 
  group_by(would.boycott, is.artist) %>% 
  summarise(n=n()) %>% 
  group_by(would.boycott) %>% 
  mutate(total=sum(n)) 

boycotts %>% 
  ggplot(aes(x=would.boycott,y=n, fill=is.artist)) +
    geom_bar(stat = "identity") +
    geom_text(aes(label=total, x=would.boycott, y=total+1)) +
    scale_fill_brewer(palette="Paired") +
    theme_minimal() +
    labs(fill=q2_varname) +
    theme(legend.position = "bottom") +
    ggtitle(q3_varname) +
    xlab(NULL) +
    ylab(NULL) 

```

Turns out, exactly half of the respondents would consider boycotting a piece of AI-generated art, if the trained-on artist was not compensated for his contribution. Some of them only if they already really liked the artist.

Maybe more surprisingly, the two respondents claiming to be artists themselves answered not to be wanting to boycott such art.

```{r, fig.width=10}
ggplot(ans, aes(y=would.boycott, fill=is.artist)) + 
  geom_bar() + 
    scale_fill_brewer(palette="Paired") +
  labs(y = NULL, x=NULL,
       title=q3_varname %>% paste( "by level of engagement with art"),
       fill=q2_varname
       ) + 
  facet_wrap(~ is.artist)
```

## Independence testing.

*Null hypothesis: If the artist the generative model was trained on were not granted any of the revenue generated by the art, other artists would be more likely to boycott that piece of art.*  

From looking at the above barplots, we cannot make out a clear pattern of who is more likely to boycott art visually.

```{r, fig.width=10}
# cut the revenue.deserved values into k levels: low to high 
table(ans$would.boycott, ans$is.artist) %>% 
  print() %>% 
  chisq.test()

# what if we replicated that data with the same results?
table(ans$would.boycott, ans$is.artist) %>% 
  chisq.test(simulate.p.value = TRUE)

```

# Research hypothesis 3

*Hypothesis: There is no statistical connection between people who would grant artists/authors more money (question 1) and people who would boycott an AI-generated piece of art, that the trained-on artist did not make any money from, meaning that respondents, while they may agree with the ethical worries of artists regarding AI-generated art, they do not act upon those values.*  

## Exploring the data

```{r, fig.width=9}
ans %>%   
  ggplot(aes(x=revenue.deserved, y=would.boycott)) +
    geom_boxplot() +
    theme_minimal() +
    ylab(NULL) +
    stat_summary(fun.y=mean, geom="point", shape=20, size=5, color="red") +
    ggtitle(q1_varname %>% paste("\nby", q3_varname)) +
    xlab(q1_varname)

```

Looking at the parallel boxplots, we can make out a difference in distribution for revenue percentages credited to trained-on artists for the boycotters and non-boycotters. The median and mean (red dots) values for the respondents willing to boycott such art are higher than for the ones not willing to. The respondents who would not boycott art created from unpaid training is all over, while the respondents who would boycott it, appear to be credit the artists more of the generated revenue.  
  
A different perspective on the answer options would be, that perhaps the difference in distributions is only there, because of the separation of "Yes" answers. Regardless, when joining the answers, "Yes" and "Only if I really like the artist", the overall distribution of the answers in regards of granted revenue looks not so similar.  

```{r}
ans$would.boycott.2 <- ans$would.boycott %>% fct_lump_min(16, other_level="Yes") 

ans %>%   
ggplot(aes(x=revenue.deserved, y=would.boycott.2)) +
  geom_boxplot() +
  theme_minimal() +
  stat_summary(fun.y=mean, geom="point", shape=20, size=5, color="red") +
  ylab(NULL) +
  ggtitle(q1_varname %>% paste("\nby", q3_varname)) +
  xlab(q1_varname)

```

The differences in mean are as follows:  

```{r, echo=FALSE}
ans %>% 
  group_by(would.boycott) %>% 
  summarise(rev_mean = mean(revenue.deserved, na.rm = T)) %>% 
  rename(`Mean credited revenue`=rev_mean) 

ans %>% 
  group_by(would.boycott.2) %>% 
  summarise(rev_mean = mean(revenue.deserved, na.rm = T)) %>% 
  rename(`Mean credited revenue`=rev_mean)

```

## Testing the differences in means

*Null-Hypothesis: There is no statistical difference in the means of the percentage of revenue credited to artists as comparing the groupy of respondents willing and not willing to boycott art, whose generative-model was trained on said artists.*

Starting out by testing the difference in means of the two groups:\
Respondents that would boycott and those that would not.

The response values are not normally distributed, therefore we use a non-parametric test, namely a Wilcoxon test.

```{r}
wilcox.test(
  filter(ans, would.boycott.2=="Yes")$revenue.deserved, 
  filter(ans, would.boycott.2=="No")$revenue.deserved, 
  paired = TRUE, 
  alternative = "two.sided",
  exact = FALSE
  )
```

The test yields a p-value of \<0.033, which lets us reject the Null-Hypothesis that there is no difference in means of granted revenue percentage between the groups.\
We can therefore confirm that for our sample, the respondents that are inclined to boycott a piece of art, if the original, trained-on artist did not get compensated, would also grant them a higher portion of the revenue that would have been generated from said piece of art.

## Exploring the non-boycotting group further

The results is not surprising, we in fact were trying to make out, whether respondents that would grant the artists a greater portion of the generated revenue would even consider boycotting the AI-generated piece, determining if the ethical values would be even acted upon.

```{r}
non.boycotter <- ans %>% 
  filter(would.boycott == "No") 

p1 <- non.boycotter %>% 
  ggplot(aes(x=revenue.deserved)) +
  geom_boxplot()
p2 <- non.boycotter %>% 
  ggplot(aes(x=revenue.deserved)) +
  geom_histogram()
```


```{r, echo=F, fig.width=6}
p1 %>% print()
p2 %>% print()

```

  

The non-boycotters' percentages of revenue granted to trained-on artists, is quite evenly distributed. Seeing as we are only looking at a total of 19 respondents, we are far from being able to make a generalizing statement to answer the original question.
